---
title: "Using fingertipsR to analyse local health data"
output: html_notebook
---

Local health data has recently been made accessible through the [Fingertips API](https://fingertips.phe.org.uk/api). This short note shows how to access the data using `fingertipsR`. We also show how to combine these data with the recent data on fast food outlet density at ward level and compare this with levels of obesity and deprivation.

```{r}
library(pacman)

p_load(tidyverse, readxl,  fingertipsR, caret, caretEnsemble, data.table, ggridges, leaps, rpart, partykit, DALEX, janitor)

```


To get the data from Fingertips we need to know that the ProfileID = 143 aand AreaTypeId = 8 for wards. We use the `fingertips_data` function to extract the data (it will take a few minutes and the download is 129 MB).

```{r get data, cache=TRUE}

lh <- fingertips_data(ProfileID = 143, AreaTypeID = 8)

```

We can obtain the fast food data from .GOV.UK as follows. The *ward* data is on the third sheet.

```{r}

url <- "https://www.gov.uk/government/uploads/system/uploads/attachment_data/file/720059/FastFoodMetadata_LA_Ward.xlsx"


fastfood <- downloader::download(url, "fastfood.xlsx")

##
setwd("~/Downloads")

fastfoodsheets <- excel_sheets("fastfood.xlsx")

fastfoodxl <- read_excel("fastfood.xlsx", "Ward Data", skip = 2) %>% janitor::clean_names()




```

We can convert the data from "long" forma to "wide" format so that each column is a variable. Because the dataset is large, we use the `data.table` package to process for speed and reduce the number of variables in the dataset.

```{r}

lh <- data.table(lh)

lh1 <- lh %>% clean_names()

lh1 <- lh1[, index := paste(indicator_id, indicator_name, timeperiod, sex) ] ## create a unique indicator variable name


lh1 <- lh1[, .(index, area_name, area_code, value)] ## reduce number of columns


lhw <- lh1 %>% spread(index, value) %>% janitor::clean_names() ## convert to wide format


names(lhw)[1:5]  


```

This gives a table of `r nrow(lhw)` rows (individual wards) and `r ncol(lhw)` variables. 

We can join now the datasets using the ward codes.

```{r}

lhw1 <- lhw %>%
  left_join(fastfoodxl, by = c("area_code" = "x2016_ward_code")) %>% 
  mutate(density = 1000 *total_fast_food_outlets/ x93225_total_population_2015_persons)

## summary stats

a <- lhw1 %>% 
  summarise_at(vars(3:75), funs(n = n(), 
               meanna = mean(is.na(.), na.rm = TRUE), 
               sd =  sd(., na.rm = TRUE))) 

a[1]




```

